---
title: "Class Test 1 Marking Scheme"
output:
  pdf_document:
          latex_engine: xelatex
          number_sections: no
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
```

```{r, eval = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(MASS)
library(readr)
library(tidyr)
```

```{r, eval = TRUE, warning = FALSE, message = FALSE}
PGA <- read_csv("PGALPGA2008.csv")
PGA$Gender <- as.factor(PGA$Gender)
```

Successful upload of `.pdf` file.

\begin{flushright}2 MARKS\end{flushright}

# Report

## Introduction {#sec:intro}

Introduction to the data being analysed and to the question of interest. No marks for copying the data description as given. 1 mark removed if the document title has not been changed.

\begin{flushright}2 MARKS\end{flushright}

## Exploratory data analysis {#sec:eda}

Summary statistics of the data with appropriate comments. 1 mark removed if the output is simply 'copy-pasted' from R.

\begin{flushright}3 MARKS\end{flushright}

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
PGA %>%
  group_by(Gender) %>%
   summarize('Accuracy (Mean)' = mean(Accuracy), 'Accuracy (Median)' = median(Accuracy),
             'Accuracy (sd)' = sd(Accuracy)) %>%
  kable(caption = '\\label{tab:summary} Mean, median and standard deviation (sd) of accuracy by gender.', booktabs = TRUE, digits = 2) %>%
  kable_styling(font_size = 9, latex_options = "hold_position")
```

<!-- This code commented out below is now deprecated and may not work -->
<!-- # ```{r, eval = TRUE, warning = FALSE, message = FALSE} -->
<!-- # skim_with(numeric = list(hist = NULL, missing = NULL, complete = NULL, n = NULL, variable = NULL)) -->
<!-- # PGA %>% -->
<!-- #   group_by(Gender) %>% -->
<!-- #   skim_to_list() %>% -->
<!-- #   .$numeric %>% -->
<!-- #   kable(col.names = c("Gender", "Variable", "Mean", "SD", "Minimum", "1st quartile", "Median", -->
<!-- #                       "3rd quartile", "Maximum"),  -->
<!-- #         caption = '\\label{tab:summary} Summary statistics on driving accuracy by gender.', -->
<!-- #         booktabs = TRUE) %>% -->
<!-- #   kable_styling(font_size = 10, latex_options = "hold_position") -->
<!-- # ``` -->

Boxplot of accuracy by gender. 1 mark removed if the plot is not appropriately labelled, and axis labels not adjusted accordingly.

\begin{flushright}2 MARKS\end{flushright}

```{r, eval = TRUE, out.width = '68%', fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:box} Driving accuracy between PGA and LPGA Tour golfers in 2008.", fig.pos = 'H'}
ggplot(PGA, aes(x = Gender, y = Accuracy)) +
  geom_boxplot() +
  labs(x = "Gender", y = "Drive accuracy (%)", title = "")
```

Comments on the boxplot related to the question of interest. 


## Formal data analysis {#sec:fda}

State the linear regression model being fitted, i.e.

$$\widehat{\mbox{accuracy}} = \widehat{\alpha} + \widehat{\beta}_{\mbox{Male}} \cdot \mathbb{I}_{\mbox{Male}}(x) $$

where

  * the intercept $\widehat{\alpha}$ is the mean accuracy for the baseline category (females);
  * $\widehat{\beta}_{\mbox{Male}}$ is the difference in the mean accuracy of males relative to the baseline category (females); and
  * $\mathbb{I}_{\mbox{Male}}(x)$ is an indicator function such that

    $$\mathbb{I}_{\mbox{Male}}(x)=\left\{
                \begin{array}{ll}
                  1 ~~~ \mbox{if the gender of} ~ x \mbox{th observation is Male},\\
                  0 ~~~ \mbox{Otherwise}.\\
                \end{array}
              \right.$$

\begin{flushright}2 MARKS\end{flushright}

```{r, eval = TRUE, warning = FALSE, message = FALSE}
model <- lm(Accuracy ~ Gender, data = PGA)
Coefs <- round(coef(model), 3)
```

Regression model output. 1 mark removed if the regression output is simply 'copy-pasted' from R.

\begin{flushright}2 MARKS\end{flushright}

```{r, eval = TRUE, warning = FALSE, message = FALSE}
get_regression_table(model)[, 1:2] %>%
  kable(caption = '\\label{tab:reg} Estimates of the intercept and slope from the fitted linear regression model.',
        booktabs = TRUE) %>%
  kable_styling(latex_options = 'HOLD_position')
```

Appropriate comments on the regression coefficients and differences between males and females.

\begin{flushright}2 MARKS\end{flushright}

```{r, eval = TRUE, warning = FALSE, message = FALSE}
regression.points <- get_regression_points(model)
```

Plots for checking model assumptions. 1 mark removed if not properly labelled. 

\begin{flushright}3 MARKS\end{flushright}

```{r, eval = TRUE, fig.width = 13, fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:resids} Scatterplots of the residuals by gender (left) and a histogram of the residuals (right).", fig.pos = 'H', message = FALSE}
p1 <- ggplot(regression.points, aes(x = Gender, y = residual)) +
      geom_jitter(width = 0.1) + 
      labs(x = "Gender", y = "Residual") +
      geom_hline(yintercept = 0, col = "blue")

p2 <- ggplot(regression.points, aes(x = residual)) +
      geom_histogram(color = "white") +
      labs(x = "Residual") 

grid.arrange(p1, p2, ncol = 2)
```

Appropriate comments on the model assumptions. 

\begin{flushright}3 MARKS\end{flushright}

## Conclusions {#sec:con}

Overall conclusions with an answer to the question of interest. 

\begin{flushright}2 MARKS\end{flushright}

General report layout. This should include figure and table captions, with marks not awarded if these are
not used. 1 mark removed if hyperlinks for sections and figures not implemented (tables are allowed no hyperlinks). 1 mark removed for consistently poor spelling mistakes/errors.

\begin{flushright}2 MARKS\end{flushright}

***
\begin{flushright}Total: 25 MARKS\end{flushright}

# Further Question 1

```{r, warning=FALSE, message=FALSE}
q1data <- read_csv("test1.csv")
```

Modify the data into `tidy` format and set the categorical variable to type `factor` using:

```{r, echo = TRUE}
q1datatidy <- gather(data = q1data, 
                    key = Fertilizer,
                    value = Yield) 
q1datatidy$Fertilizer <- as.factor(q1datatidy$Fertilizer)
```

2 marks are awarded for converting to the `tidy` format, while 1 mark is awarded for converting `Fertilizer` to a factor.

\begin{flushright}3 MARKS\end{flushright}

To graphically compare the distribution of the yields from the four fertilizers we produce boxplots.

```{r, echo = TRUE, fig.cap = "\\label{fig:box1} Distribtuion of the crop yields for each fertilizer (A-C) and control group (D).", fig.pos = "H", out.width = '60%', fig.align = "center"}
ggplot(data = q1datatidy, aes(y = Yield, x = Fertilizer)) +
         geom_boxplot()
```

2 marks are awarded for producing the appropriately labelled boxplot.

The boxplots suggest that there are **two outliers** (i.e. extremely small yields compared to the other 8 yields) for fertilizer A.  These should be at the very least investigated further and potentially removed from the data. The boxplots also suggest fertilizer C produces the **biggest crop** yields, then B, then A (with **large variation**) with the control group (D) having the smallest yields.

2 marks are awarded for appropriate comments relating to the data and boxplots.

\begin{flushright}4 MARKS\end{flushright}

***
\begin{flushright}Total: 7 MARKS\end{flushright}

# Further Question 2

```{r, echo = TRUE, eval = TRUE, warning = FALSE}
set.seed(10)
n_sim <- 100
corr <- 0.6
mu <- c(10, 18)
sigma <- matrix(c(1, corr, corr, 1), 2, 2)
sim <- mvrnorm(n_sim, mu = mu, Sigma = sigma)
colnames(sim) <- c("X", "Y")
```

2 marks are awarded for correctly identifying the number of observations, the means of $X$ and $Y$, and the correlation matrix. An additional 2 marks are awarded for the correct use of the `mvrnorm()` function.

\begin{flushright}4 MARKS\end{flushright}

```{r, echo = TRUE, eval = TRUE, warning = FALSE, out.width = '60%', fig.align = "center", fig.cap = "Scatterplot of Y against X.", fig.pos = "H"}
sim <- as.data.frame(sim)
ggplot(data = sim, aes(x = X, y = Y)) +
  geom_point() +
  labs(x = "X", y = "Y")

sim %>%
  summarize(cor(X, Y))
```

1 mark for producing the scatterplot and comments, and 1 mark for obtaining the correlation coefficient.

\begin{flushright}2 MARKS\end{flushright}

***
\begin{flushright}Total: 6 MARKS\end{flushright}

***
\begin{flushright}Total: 40 MARKS\end{flushright}