---
title: "report"
author: "Data Analysis"
output: 
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
fig_caption: yes
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE, warning = FALSE)
```

# Introduction{#sec:intro}
This is an example

## first{#sec:sub1}
Well done,bro

## second{#sec:sub2}
keep going 


# Code chunk{#sec:chunk}


## show output only{#sec:sub11}
```{r echo=FALSE}
print(1:5)
```

## show code and output{#sec:sub22}
```{r echo=TRUE}
print(6:10)
```

## show nothing(run code){#sec:sub33}
```{r include=FALSE}
print(6:10)
```

## show nothing(don't run code){#sec:sub44}
```{r eval=FALSE, include=FALSE}
print(7:19)
```

# Inline code{#sec:Inline}

`r sin(pi/2)` 


`r 2+2`



# Table{#sec:table}

```{r}
x <- 1:10
y <- x^2
lmr <- lm(y ~ x)
co <- summary(lmr)$coefficients
knitr::kable(co)
```

```{r table}
knitr::kable(head(iris,n=5),caption='\\label{tab:iris} The first 5 rows')
```

# Mathematics{#sec:maths}

$$y_i = \alpha + \beta x_i + \epsilon_i, ~~~~ \epsilon_i \sim N(0, \sigma^2),$$


$$y_i = \alpha + \beta_{\mbox{Male}} \cdot \mathbb{I}_{\mbox{Male}}(x),$$


# Figures{#sec:figures}
```{r include=FALSE}

library(ggplot2)
library(nycflights13)
```

Figure \ref{fig:scat} displays a scatterplot of the teaching score against beauty score. Here, there appears to be a positive relationship between teaching and beauty scores. Hence, teaching score tends to increase with beauty score. However, as seen from Figure \ref{fig:scat} and the observed correlation coefficient , it is only a weakly positive relationship. A linear regression model will now be fitted to assess the relationship between teaching and beauty scores.
\newpage

```{r eval=TRUE,fig.align = "center", fig.cap = "\\label{fig:scat} Relationship between teaching and beauty scores. The best-fitting line has been superimposed.",fig.pos="h"}
Alaska<-flights[flights$carrier=="AS",]
ggplot(data=Alaska,mapping=aes(x=dep_delay,y=arr_delay))+
  geom_point()+
  labs(x="Depature delay(minutes)",y="Arrival delay(minutes)",title="Alaska Airlines flights leaving NYC in 2013")
```



## inside code{#sec:sub111}
```{r eval=TRUE, fig.align="center", fig.cap="try", fig.width=6, out.width='80%',fig.pos="h"}
x=1:100
y=0.5*x
plot(x,y)
```







