---
title: "Class Test 1 Marking Scheme"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: no
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
```

```{r, eval = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(MASS)
library(readr)
library(tidyr)
```

```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
PGA <- read_csv("PGA2008.csv")
```

Successful upload of `.pdf` file.

\begin{flushright}2 MARKS\end{flushright}

# Report

## Introduction {#sec:intro}

Introduction to the data being analysed and to the question of interest. No marks for copying the data description as given. 1 mark removed if the document title has not been changed.

\begin{flushright}2 MARKS\end{flushright}

## Exploratory data analysis {#sec:eda}

Summary statistics on driving accuracy and distance with appropriate comments. One mark removed if the output is simply 'copy-pasted' from R.

\begin{flushright}3 MARKS\end{flushright}

```{r, eval = TRUE}
my_skim <- skim_with(numeric = sfl(hist = NULL))
my_skim(PGA)  %>%
  dplyr::select(-skim_type) %>%
  kable(col.names = c("Variable", "Missing", "Complete", "Mean", "SD", "Min.", "1st Q.", "Median",
                        "3rd Q.", "Max."),
        caption = '\\label{tab:summary} Summary statistics on driving accuracy and distance.',
        booktabs = TRUE, format = "latex", digits = 2) %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```

```{r cor, echo = FALSE, eval = FALSE}
PGA %>%
  get_correlation(formula = Accuracy ~ Distance)
```

Scatterplot of drive accuracy against distance. One mark removed if the plot is not appropriately labelled, and axis labels not adjusted accordingly.

\begin{flushright}2 MARKS\end{flushright}

The regression line can be superimposed here or within the formal data analysis section.

\begin{flushright}1 MARK\end{flushright}

```{r scatplot, echo = FALSE, eval = TRUE, out.width = '68%', fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:scat} Relationship between driving accuracy and distance. The best-fitting line has been superimposed.", fig.pos = 'H', message = FALSE}
ggplot(PGA, aes(x = Distance, y = Accuracy)) +
  geom_point() +
  labs(x = "Distance (yards)", y = "Accuracy (%)",
       title = "Relationship between drive distance and accuracy in the PGA 2008") +
  geom_smooth(method = "lm", se = FALSE)
```

Comments on the scatterplot related to the question of interest. 

\begin{flushright}2 MARKS\end{flushright}

## Formal data analysis {#sec:fda}

State the linear regression model being fitted. The linear regression model that will be fitted to the data is as follows:

$$y_i = \alpha + \beta x_i + \epsilon_i, ~~~~ \epsilon_i \sim N(0, \sigma^2),$$

where $y_i$ and $x_i$ denote the **accuracy** and **distance** of the $i^{th}$ golfer, respectively. The intercept is denoted by $\alpha$, while the slope is given by $\beta$. The $i^{th}$ random error component is denoted by $\epsilon_i$ and are normally distributed with mean zero and variance $\sigma^2$.

\begin{flushright}1 MARK\end{flushright}

```{r coefs, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
model <- lm(Accuracy ~ Distance, data = PGA)
Coefs <- round(coef(model), 3)
```

Regression model output. One mark removed if the regression output is simply 'copy-pasted' from R.

\begin{flushright}2 MARKS\end{flushright}

```{r regtable, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
get_regression_table(model)[, 1:2] %>%
  kable(caption = '\\label{tab:reg} Estimates of the intercept and slope from the fitted linear regression model.',
        booktabs = TRUE) %>%
  kable_styling(latex_options = 'HOLD_position')
```

Appropriate comments on the regression coefficients and the relationship between drive accuracy and distance.

\begin{flushright}2 MARKS\end{flushright}

```{r regpoints, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
regression.points <- get_regression_points(model)
```

Plots for checking model assumptions. One mark removed if not properly labelled. 

\begin{flushright}3 MARKS\end{flushright}

```{r residplots, echo = FALSE, eval = TRUE, fig.width = 13, fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:resids} Scatterplots of the residuals against distance (left) and the fitted values (right).", fig.pos = 'H'}
p1 <- ggplot(regression.points, aes(x = Distance, y = residual)) +
      geom_point() +
      labs(x = "Drive distance (in yards)", y = "Residual") +
      geom_hline(yintercept = 0, col = "blue", size = 1) 

p2 <- ggplot(regression.points, aes(x = Accuracy_hat, y = residual)) +
      geom_point() +
      labs(x = "Fitted values", y = "Residual") +
      geom_hline(yintercept = 0, col = "blue", size = 1) 

grid.arrange(p1, p2, ncol = 2)
```

```{r residplots2, echo = FALSE, eval = TRUE, out.width = '58%', fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:residhist} Histogram of the residuals.", message = FALSE, fig.pos = "H"}
ggplot(regression.points, aes(x = residual)) +
      geom_histogram(color = "white") +
      labs(x = "Residual") 
```

Appropriate comments on the model assumptions. 

\begin{flushright}3 MARKS\end{flushright}

## Conclusions {#sec:con}

Overall conclusions with an answer to the question of interest. 

\begin{flushright}2 MARKS\end{flushright}

General report layout. This should include figure and table captions, with marks not awarded if these are
not used. One mark removed if hyperlinks for sections and Figures not implemented (Tables are allowed no hyperlinks). 

\begin{flushright}2 MARKS\end{flushright}

***
\begin{flushright}Total: 25 MARKS\end{flushright}

# Further Question 1

```{r, warning=FALSE, message=FALSE, echo = TRUE}
q1data <- read_csv("EPL.csv")
```

Modify the data into `tidy` format and set the categorical variable to type `factor` using:

```{r, echo = TRUE}
q1datatidy <- gather(data = q1data,
                    key = Year,
                    value = Points,
                    - Team)
q1datatidy$Team <- as.factor(q1datatidy$Team)
```

2 marks are awarded for converting to the `tidy` format, while 1 mark is awarded for converting `Team` to a factor.

\begin{flushright}3 MARKS\end{flushright}

To graphically compare the point distributions of the six teams we produce boxplots.

```{r, echo = TRUE, fig.cap = "\\label{fig:box1} Distribtuion of the total points for each team.", fig.pos = "H", out.width = '60%', fig.align = "center"}
ggplot(data = q1datatidy, aes(y = Points, x = Team)) +
         geom_boxplot() +
         scale_x_discrete(labels = c("Arsenal", "Chelsea", "Everton", "Liverpool",
                                      "Man United", "Spurs"))
```

2 marks are awarded for producing the appropriately labelled boxplot.

2 marks are awarded for appropriate comments relating to the data and boxplots.

\begin{flushright}4 MARKS\end{flushright}

***
\begin{flushright}Total: 7 MARKS\end{flushright}

# Further Question 2

```{r, echo = TRUE, eval = TRUE, warning = FALSE}
set.seed(10)
n_sim <- 100
corr <- -0.65
mu <- c(12, 19)
VarX <- 2
VarY <- 1
sqrt.var <- sqrt(prod(c(VarX, VarY)))
sqrt.var.mat <- matrix(c(1, sqrt.var, sqrt.var, 1), 2, 2)
sigma <- matrix(c(VarX, corr, corr, VarY), 2, 2) * sqrt.var.mat
sim <- mvrnorm(n_sim, mu = mu, Sigma = sigma)
colnames(sim) <- c("X", "Y")
```

2 marks are awarded for correctly identifying the number of observations, the means of $X$ and $Y$, and the covariance matrix. An additional 2 marks are awarded for the correct use of the `mvrnorm()` function.

\begin{flushright}4 MARKS\end{flushright}

```{r, echo = TRUE, eval = TRUE, warning = FALSE, out.width = '60%', fig.align = "center", fig.cap = "Scatterplot of Y against X.", fig.pos = "H"}
sim <- as.data.frame(sim)
ggplot(data = sim, aes(x = X, y = Y)) +
  geom_point() +
  labs(x = "X", y = "Y")

sim %>%
  summarize(cor(X, Y))
```

1 mark for producing the scatterplot and comments, and 1 mark for obtaining the correlation coefficient.

\begin{flushright}2 MARKS\end{flushright}

***
\begin{flushright}Total: 6 MARKS\end{flushright}

***
\begin{flushright}Total: 40 MARKS\end{flushright}